---
title: "v18 QC Fusion files"
output: html_notebook
params:
  histology:
    label: "Clinical file" 
    value: release-v19-20210112/pbta-histologies-base.tsv
    input: file
  oldDataFolder:
    label: "Old release folder soft linked to this directory"
    value: release-v18-20201123
    input: file
  newDataFolder:
    label: "Input new rsem files"
    value: output
    input: file
  outputFolder:
    label: "Output release folder"
    value: release-v19-20210112
    input: string  
---


This markdown helps with QC for merging new RNAseq rsem.gene files.

     Note v19 has the following changes:
     
     Stranded:
       BS_JXF8A2A6 , the reason behind the removal is in  [issue](https://github.com/AlexsLemonade/OpenPBTA-analysis/issues/862)

     
##### Checks

1. PRINTS: dimentions of new and old datasets
2. CHECKS: sample overlap between new and histology file
3. If all above checks pass then final "Checks PASSED/Checks FAILED" is printed


```{r Read input}
if(!require("readr")){
    install.packages("readr")
}
if(!require("tidyverse")){
    install.packages("tidyverse")
}
if(!require("reshape2")){
    install.packages("reshape2")
}
if(!require("qdapRegex")){
    install.packages("qdapRegex")
}
root_dir <- rprojroot::find_root(rprojroot::has_dir(".git"))
#load required packages
suppressPackageStartupMessages(library("readr"))
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("reshape2"))
suppressPackageStartupMessages(library("qdapRegex"))

```

### Input

```{r}
# read released fusion file
oldpbtaStarfusion<-read_tsv(file.path(params$oldDataFolder,"pbta-fusion-starfusion.tsv.gz"))
oldpbtaArriba<-read_tsv(file.path( params$oldDataFolder,"pbta-fusion-arriba.tsv.gz"))

# read newly created fusion file
newpbtaStarfusion<- oldpbtaStarfusion %>%
  filter(tumor_id != "BS_JXF8A2A6")

newpbtaArriba<- oldpbtaArriba %>%
  filter(tumor_id != "BS_JXF8A2A6")

# read histology file
# histology file
clinical<-read.delim(file.path(params$histology), stringsAsFactors = FALSE)
clinical_rna<-clinical %>% dplyr::filter(experimental_strategy=="RNA-Seq") 

# results folder
outputFolder<-params$outputFolder

```

#### v18 fuson files

```{r Dimension of old data }
#old starfusion
dim(oldpbtaStarfusion)
#old arriba
dim(oldpbtaArriba)


```

#### New fusion files

```{r Dimension of new data }
#old starfusion
dim(newpbtaStarfusion)
#old arriba
dim(newpbtaArriba)


```


```{r Check overlap of samples from histology file }
# check samples in starfusion file
fusedSampleCheck1<-setdiff(newpbtaStarfusion$tumor_id,clinical$Kids_First_Biospecimen_ID)

# check samples in arriba fusion file
fusedSampleCheck2<-setdiff(newpbtaArriba$tumor_id,clinical$Kids_First_Biospecimen_ID)

if (all(is_empty(c(fusedSampleCheck1,fusedSampleCheck2)))){
  print ("Samples in fusion calls overlap histology file ")
  sampleCountCheck<-TRUE
} else{
  stop ("Different sample ids found! Check samples in new files")
  sampleCountCheck<-FALSE
}

```


```{r  Check all Checks and save RDS}
if(all(sampleCountCheck)){
  print("Checks PASSED, coping file to release folder")
  write_tsv(newpbtaStarfusion,file.path(outputFolder,"pbta-fusion-starfusion.tsv.gz"))
  write_tsv(newpbtaArriba,file.path(outputFolder,"pbta-fusion-arriba.tsv.gz"))
  
} else{
  stop("Checks FAILED. Please REVIEW")
}
```







