---
title: "Histologies File QC"
data: 2021-01-27 
output:
  pdf_document:
    latex_engine: xelatex
author: "Jo Lynne Rokita, Krutika Gaonkar (D3B)"
params:
  latest_histology:
    label: "latest pull of base histology"
    value: 20210126-data/ADAPT-base/pbta_histologies.csv
    input: file
  prev_histology:
    label: "previous release of base histology"
    value: 20201215-data/pbta-histologies.tsv
    input: file  
  output:
    label: "release folder name"
    value: "20210126-data"
    input: string
  add_ids:
    label: "add Kids_First_Biospecimen_IDs from base not in previous release"
    value: ""
    input: string
  remove_ids:
    label: "remove Kids_First_Biospecimen_IDs from previous release in ouput"
    value: ""
    input: string   
toc: TRUE
toc_float: TRUE
editor_options: 
  chunk_output_type: inline
---

In this notebook we are using v18 base histology to create a base histology for v19 release. "Base histology" file has the basic clinical information manifest that is required by subtyping modules to add in OpenPBTA subtyping information.

The v18 base histologies was generated in this script: [script](https://github.com/d3b-center/D3b-codes/blob/master/OpenPBTA_v18_release_QC/QC_histology_v18.Rmd).

CNS_region values were mis-assigned by a bug in v18 which will be fixed and QC-ed as well [#14](https://github.com/d3b-center/D3b-codes/issues/14) and original issue on OpenPBTA is in [838](https://github.com/AlexsLemonade/OpenPBTA-analysis/issues/838)


```{r global_options, include=F}
knitr::opts_chunk$set(error = F, echo = T, warning = T, message = T)

```


# Load packages
```{r}
suppressMessages(library(emo))
suppressMessages(library(tidyverse))

```

# Directories and Files
## Directories
```{r}
# Input directory
input_dir <- file.path("input")
# soft linked previous release histology
prev_hist_file <- params$prev_histology

# adapt histology
latest_hist_file <- params$latest_histology

##--- KEEP LINK to G-DRIVE --- ##

# pathology diagnosis is needed to match tumor samples 
# to broad/short histology
#path_dx <- read_sheet('https://docs.google.com/spreadsheets/d/1fDXt_YODcSAWDvyI5ISBVhUCu4b5-TFCVWMOwiPeMwk/edit#gid=0',range="pathology_diagnosis_for_subtyping") %>%
#  dplyr::select(pathology_diagnosis,broad_histology, short_histology) %>%
#  write_tsv(file.path(input_dir,"pathology_diagnosis_for_subtyping.tsv"))

# pathology free text diagnosis is needed to match to 
# samples marked as "Other" in pathology_diagnosis
#path_free_text <- read_sheet('https://docs.google.com/spreadsheets/d/1fDXt_YODcSAWDvyI5ISBVhUCu4b5-TFCVWMOwiPeMwk/edit#gid=0',range="pathology_free_text_diagnosis_for_subtyping") %>%
#  dplyr::select(pathology_free_text_diagnosis,broad_histology, short_histology)%>%
#  write_tsv(file.path(input_dir,"pathology_free_text_diagnosis_for_subtyping.tsv"))

## ----------- ##


```


### Read in old base histology

```{r}
prev_hist <- read_tsv(prev_hist_file,
                      # NAs are being read as logical so specifying as character here
                           col_types = readr::cols(molecular_subtype = readr::col_character(),
                                                   short_histology = readr::col_character(),
                                                   integrated_diagnosis = readr::col_character(),
                           broad_histology = readr::col_character(),
                           Notes = readr::col_character()))

path_dx <- read_tsv(file.path(input_dir,"pathology_diagnosis_for_subtyping.tsv")) %>%
  dplyr::select(pathology_diagnosis, broad_histology, short_histology)
path_free_text <- read_tsv(file.path(input_dir,"pathology_free_text_diagnosis_for_subtyping.tsv")) %>%
  dplyr::select(pathology_free_text_diagnosis, broad_histology, short_histology)

```

### Subset new file for only those sampleIDs required
v18 but we will remove BS_JXF8A2A6 for v19 [#862](https://github.com/AlexsLemonade/OpenPBTA-analysis/issues/862)
```{r}
latest_hist <- read_tsv(latest_hist_file,
                      # NAs are being read as logical so specifying as character here
                           col_types = readr::cols(molecular_subtype = readr::col_character(),
                                                   short_histology = readr::col_character(),
                                                   integrated_diagnosis = readr::col_character(),
                           broad_histology = readr::col_character(),
                           Notes = readr::col_character()))

# get ids to subset 
id_to_subset <- prev_hist %>%
  pull(Kids_First_Biospecimen_ID)
```

## Add ids to previous release?

```{r}
if (params$add_ids != ""){
  add_ids <- unlist(str_split(params$add_ids, ","))
  # add new ids to previous releases
  id_to_subset <- c( id_to_subset, add_ids)
  print(paste(toString(add_ids)," added"))
}

```


### subset to previous ids ( and new ids if provided)
```{r}
# subset final histology
latest_hist <- latest_hist %>%
  filter(Kids_First_Biospecimen_ID  %in% id_to_subset)


```

## Check 1: Assess dimensions whether new column names match the old 

### Check 1a: assess ids overlap in new and old `r emo::ji("x")`
```{r}
source("code/util/check_rows_cols.R")
check_rows(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "Kids_First_Biospecimen_ID")
```


### Check 1b: assess columns overlap in new and old `r emo::ji("white_check_mark")`
```{r}
check_cols(new_hist = latest_hist,old_hist = prev_hist)

```

## Check 2: Assess levels of histology columns `r emo::ji("x")`

### Check 2a: path_dx and path_free_text_dx is used to match later so should have the same values in new histology

```{r}

check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "pathology_diagnosis",output_dir = params$output)

```
```{r}

check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "pathology_free_text_diagnosis",output_dir = params$output)
```

### Check 2b: Normals, these should not have path_dx, int_dx,molecular_subtype, broad/short_hist `r emo::ji("x")`
```{r}
latest_hist_normals <- latest_hist %>% 
  filter(sample_type=="Normal")
prev_hist_normals <- prev_hist %>%
  filter(sample_type=="Normal")

key_column_name = c("pathology_free_text_diagnosis","pathology_diagnosis","primary_site")

distinct(prev_hist_normals[,key_column_name])
distinct(latest_hist_normals[,key_column_name])

```

## Check3 tables per column changes

### Check 3a Experimental strategy `r emo::ji("x")`

```{r}

check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "experimental_strategy",output_dir = params$output)

```

### Check 3b Sample Type  `r emo::ji("x")`
```{r}

check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "sample_type",output_dir = params$output)


```


### Check 3c Tumor Descriptor `r emo::ji("x")`

```{r}

check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "tumor_descriptor",output_dir = params$output)

```

### Check 3d Composition  `r emo::ji("x")`
```{r}
# update composition with to match new terms to previous composition terms


check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "composition",params$output)

```

### Check 3f RNA library `r emo::ji("x")`
```{r}

check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "RNA_library",output_dir = params$output)

```

### Check 3g: Cohort `r emo::ji("white_check_mark")`
```{r}
check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "cohort",output_dir = params$output)
```

### Check 3h: Sample and aliquot IDs - any changes? `r emo::ji("x")`
```{r}

check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "sample_id",output_dir = params$output)


```

```{r}
check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "aliquot_id", output_dir = params$output)

```

### Check 3i: Sequencing Center `r emo::ji("x")`
```{r}

check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "seq_center",output_dir = params$output)

```

### Check 3f: primary_site `r emo::ji("x")`
```{r}

check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "primary_site",output_dir = params$output)

```

## Update CNS_region 

json file was generated from the CNS_region updates ticket [838](https://github.com/AlexsLemonade/OpenPBTA-analysis/issues/838). 

#### Match CNS_region matching primary_site and update

```{r}
source("code/util/primary_site_matched_CNS_region.R")
latest_hist <- get_CNS_region(histology=latest_hist,
                   CNS_match_json=file.path(input_dir,"CNS_primary_site_match.json")) 

```


Which samples had different CNS_region in v18?
```{r}

diff_cns <-latest_hist %>%
  left_join(prev_hist[,c("Kids_First_Biospecimen_ID","CNS_region")],by=c("Kids_First_Biospecimen_ID"), suffix = c(".v19",".v18")) %>%
    dplyr::select(Kids_First_Biospecimen_ID,CNS_region.v18,CNS_region.v19,primary_site) %>%
  dplyr::filter(CNS_region.v18 != CNS_region.v19)


diff_cns 
```

```{r}
diff_cns$CNS_region.v19 %>% table()
```

135 samples were incorrectly assigned CNS_region in v18. 129 on these should be 'Mixed' and 6 'Other', fixed with an updated `cns_region_check()` function in this notebook.

## Update broad_histology and short_histology
Match by pathology_diagnosis and pathology_free_text_diagnosis (Other)
 
#### By path_free_text for "Other" diagnosed 

Only samples with 'Other' in pathology_diagnosis will be need to be matched by path_free_text

```{r}
latest_hist<- dplyr::select(latest_hist,c(-broad_histology,-short_histology))
latest_hist_other <- latest_hist  %>%
  dplyr::filter(pathology_diagnosis == "Other") %>%
  left_join(path_free_text,by="pathology_free_text_diagnosis") 
```

#### By path_dx for all tumors other than "Other"
Remove samples with 'Other' in pathology_diagnosis that was already matched above
```{r}
latest_hist <- latest_hist  %>%
  dplyr::filter(pathology_diagnosis != "Other"|
                  # add Normals
                  sample_type=="Normal") %>%
  left_join(path_dx,by="pathology_diagnosis") %>%
  bind_rows(latest_hist_other) 
```


### Check broad_histology
```{r}
check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "broad_histology",output_dir = params$output)

```



### Check short_histology
```{r}
check_values(new_hist = latest_hist,old_hist = prev_hist,
          column_name = "short_histology",output_dir = params$output)

```

## Remove ids from previous release?
```{r}

if (params$remove_ids != ""){
  remove_ids <- unlist(str_split(params$remove_ids, ","))
  # remove ids from previous releases
  latest_hist <- latest_hist %>%
    filter(!Kids_First_Biospecimen_ID %in% remove_ids)
  print(paste(toString(remove_ids)," removed"))
}


```


# Write new file
```{r}
write.table(latest_hist, file.path(params$output,"pbta-histologies-base.tsv")
            , sep = "\t", quote = F, col.names = T, row.names = F)
```


```{r}
sessionInfo()
```

