---
title: "v19 gene QC RNAseq rsem files"
output: html_notebook
params:
  histology:
    label: "histology file for v19" 
    value: release-v19-20210112/pbta-histologies-base.tsv
    input: file
  oldDataFolder:
    label: "Old release folder soft linked to this directory"
    value: release-v18-20201123
    input: file
  newDataFolder:
    label: "Input new rsem files"
    value: output
    input: file
  outputFolder:
    label: "Output release folder"
    value: release-v19-20210112
    input: string 
  polyaTPM:
    label: "Filename of polya isoform TPM files"
    value: pbta-gene-expression-rsem-tpm.polya.rds
    input: string    
  polyaCount:
    label: "Filename of polya isoform Count files"
    value: pbta-gene-counts-rsem-expected_count.polya.rds
    input: string  
  polyaFPKM: 
    label: "Filename of polya isoform FPKM files"
    value: pbta-gene-expression-rsem-fpkm.polya.rds
    input: string  
  strandedTPM:
    label: "Filename of stranded isoform TPM files"
    value: pbta-gene-expression-rsem-tpm.stranded.rds
    input: string  
  strandedCount:
    label: "Filename of stranded isoform Count files"
    value: pbta-gene-counts-rsem-expected_count.stranded.rds
    input: string  
  strandedFPKM:
    label: "Filename of stranded isoform FPKM files"
    value: pbta-gene-expression-rsem-fpkm.stranded.rds
    input: string     
---


This markdown helps with QC for merging new RNAseq rsem.gene files.

     Note v19 has the following changes:
     
     Stranded:
       BS_JXF8A2A6 , the reason behind the removal is in  [issue](https://github.com/AlexsLemonade/OpenPBTA-analysis/issues/862)

     
##### Checks

1. PRINTS: dimentions of new and old datasets
2. CHECKS: gene overlap between new and old datasets
3. CHECKS: Sample overlap with clinical files
4. CHECKS: gene order with newly formed stranded data and old polya datasets
5. If all above checks pass then final "Checks PASSED/Checks FAILED" is printed

### Setup

```{r Read input}
if(!require("readr")){
    install.packages("readr")
}
if(!require("tidyverse")){
    install.packages("tidyverse")
}
if(!require("reshape2")){
    install.packages("reshape2")
}
if(!require("qdapRegex")){
    install.packages("qdapRegex")
}

root_dir <- rprojroot::find_root(rprojroot::has_dir(".git"))
#load required packages
suppressPackageStartupMessages(library("readr"))
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("reshape2"))
suppressPackageStartupMessages(library("qdapRegex"))
```



```{r}
#read old rsem tpm files
oldStrandedTPM<-readRDS(file.path(params$oldDataFolder,params$strandedTPM))
oldPolyaTPM<-readRDS(file.path( params$oldDataFolder,params$polyaTPM))

# read new fpkm files
oldStrandedFPKM<-readRDS(file.path(params$oldDataFolder,params$strandedFPKM))
oldPolyaFPKM<-readRDS(file.path(params$oldDataFolder,params$polyaFPKM))


#read old rsem expected count files
oldStrandedCount<-readRDS(file.path(params$oldDataFolder,params$strandedCount))
oldPolyaCount<-readRDS(file.path(params$oldDataFolder,params$polyaCount))

# read new fpkm files
newStrandedFPKM<- oldStrandedFPKM %>%
  select(-BS_JXF8A2A6)

# read new strsanded tpm files
newStrandedTPM<- oldStrandedTPM %>%
  select(-BS_JXF8A2A6)

# read new stranded count files
newStrandedCount<-oldStrandedCount %>%
  select(-BS_JXF8A2A6)

# histology file
clinical<-read.delim(file.path( params$histology), stringsAsFactors = FALSE)
clinical_rna<-clinical %>% dplyr::filter(experimental_strategy=="RNA-Seq") 

# results folder
outputFolder<-params$outputFolder
```

#### v18 Stranded data

```{r Dimension of old stranded data }

#old FPKM data
dim(oldStrandedFPKM)
#old TPM data
dim(oldStrandedTPM)
#old count data"
dim(oldStrandedCount)

```

#### New stranded data

```{r Dimension of new stranded data }

#new TPM data
dim(newStrandedFPKM)
#new TPM data
dim(newStrandedTPM)
#new count data
dim(newStrandedCount)

```


```{r Check common/missing genes for stranded biospecimens}

# check with gene names in FPKM file of v17 and newly sequenced data
fpkmGeneCheck1<-setdiff(oldStrandedFPKM$gene_id,newStrandedFPKM$gene_id)
fpkmGeneCheck2<-setdiff(newStrandedFPKM$gene_id,oldStrandedFPKM$gene_id)

# check with gene names in TPM file of v17 and newly sequenced data
tpmGeneCheck1<-setdiff(oldStrandedTPM$gene_id,newStrandedTPM$gene_id)
tpmGeneCheck2<-setdiff(newStrandedTPM$gene_id,oldStrandedTPM$gene_id)

# check with gene names in count file of v17 and newly sequenced data
countGeneCheck1<-setdiff(oldStrandedCount$gene_id,newStrandedCount$gene_id)
countGeneCheck2<-setdiff(newStrandedCount$gene_id,oldStrandedCount$gene_id)

if (all(is_empty(c(countGeneCheck1,tpmGeneCheck1,countGeneCheck2,tpmGeneCheck2)))){
  print ("Common genes in both datasets ")
  geneCountCheck<-TRUE
} else{
  stop ("Different genes found! Check genes in new and old datasets")
  geneCountCheck<-FALSE
}
```


```{r Check Sample count between new dataset for stranded biospecimens and clinical file }
clinical_rna_stranded<-clinical_rna %>% filter(RNA_library=="stranded") 
#total RNAseq samples in histology
nrow(clinical_rna_stranded)

# check sample names of FPKM file in newly sequenced data and new release histologies
newSample<-colnames(newStrandedFPKM[,-1]) 
fpkmSampleCheck1<-setdiff(newSample , clinical_rna_stranded$Kids_First_Biospecimen_ID)
fpkmSampleCheck2<-setdiff( clinical_rna_stranded$Kids_First_Biospecimen_ID,newSample)


# check sample names of TPM file in newly sequenced data and new release histologies
newSample<-colnames(newStrandedTPM[,-1]) 
tpmSampleCheck1<-setdiff( newSample , clinical_rna_stranded$Kids_First_Biospecimen_ID)
tpmSampleCheck2<-setdiff( clinical_rna_stranded$Kids_First_Biospecimen_ID,newSample)

# check sample names of count file in newly sequenced data and new release histologies
newSample<-colnames(newStrandedCount[,-1]) 
countSampleCheck1<-setdiff( newSample , clinical_rna_stranded$Kids_First_Biospecimen_ID)
countSampleCheck2<-setdiff( clinical_rna_stranded$Kids_First_Biospecimen_ID,newSample)

if (all(is_empty(c(countSampleCheck1,tpmSampleCheck1,fpkmSampleCheck1,countSampleCheck2,tpmSampleCheck2,fpkmSampleCheck2)))){
  print ("New samples in rsem files completely overlap with RNAseq files in new histology file")
  SampleCountCheck<-TRUE
} else{
  stop ("Sample count inaccurate! Check Sample in new and old datasets")
  SampleCountCheck<-FALSE
}
```


```{r Check colnames for gene files for stranded biospecimens should be "gene_id" and bs_ids }

# check col names of FPKM file of v18 in newly merged file and expected output
oldSample<-colnames(oldStrandedFPKM) 
newSample<-colnames(newStrandedFPKM) 
fpkmColCheck1<-setdiff(newSample , c("gene_id",clinical_rna_stranded$Kids_First_Biospecimen_ID))
fpkmColCheck2<-setdiff( c("gene_id",clinical_rna_stranded$Kids_First_Biospecimen_ID),newSample)

# check col names of count file of v18 in newly merged file and expected output
oldSample<-colnames(oldStrandedCount) 
newSample<-colnames(newStrandedCount) 
countColCheck1<-setdiff(newSample , c("gene_id",clinical_rna_stranded$Kids_First_Biospecimen_ID))
countColCheck2<-setdiff( c("gene_id",clinical_rna_stranded$Kids_First_Biospecimen_ID),newSample)

# check col names of TPM file of v18 in newly merged file and expected output
oldSample<-colnames(oldStrandedTPM) 
newSample<-colnames(newStrandedTPM) 
tpmColCheck1<-setdiff(newSample , c("gene_id",clinical_rna_stranded$Kids_First_Biospecimen_ID))
tpmColCheck2<-setdiff( c("gene_id",clinical_rna_stranded$Kids_First_Biospecimen_ID),newSample)

if (all(is_empty(c(countColCheck1,tpmColCheck1,fpkmColCheck1,countColCheck2,tpmColCheck2,fpkmColCheck2)))){
  print ("Colnames correct in new file")
  SampleCountCheck<-TRUE
} else{
  stop ("Colnames incorrect! Check colnames in new and old datasets")
  SampleCountCheck<-FALSE
}

```




#### v18 polya data

```{r Dimension of old polya dataset}
#old TPM data
dim(oldPolyaTPM)
#old count data
dim(oldPolyaCount)
# old FPKM data
dim(oldPolyaFPKM)
```

#### New polya data
Same as v18

```{r Dimension of new polya dataset }

# new TPM data
newPolyaTPM <- oldPolyaTPM 

# new count data
newPolyaCount <- oldPolyaCount 

# new FPKM data
newPolyaFPKM <- oldPolyaFPKM

# dim new TPM data
dim(newPolyaTPM)
# dim new count data
dim(newPolyaCount)
# dim new kallisto data
dim(newPolyaFPKM)
```

```{r Check common/missing genes for polya biospecimen}

# check with gene names in FPKM file of v17 and newly sequenced data
fpkmGeneCheck1<-setdiff(oldPolyaFPKM$gene_id,newPolyaFPKM$gene_id)
fpkmGeneCheck2<-setdiff(newPolyaFPKM$gene_id,oldPolyaFPKM$gene_id)

# check with gene names in TPM file of v17 and newly sequenced data
tpmGeneCheck1<-setdiff(oldPolyaTPM$gene_id,newPolyaTPM$gene_id)
tpmGeneCheck2<-setdiff(newPolyaTPM$gene_id,oldPolyaTPM$gene_id)

# check with gene names in count file of v17 and newly sequenced data
countGeneCheck1<-setdiff(oldPolyaCount$gene_id,newPolyaCount$gene_id)
countGeneCheck2<-setdiff(newPolyaCount$gene_id,oldPolyaCount$gene_id)

if (all(is_empty(c(countGeneCheck1,tpmGeneCheck1,countGeneCheck2,tpmGeneCheck2)))){
  print ("Common genes in both datasets ")
  geneCountCheck<-TRUE
} else{
  stop ("Different genes found! Check genes in new and old datasets")
  geneCountCheck<-FALSE
}
```


```{r Check Sample count between new dataset for polya biospecimen and clinical file }
clinical_rna_Polya<-clinical_rna %>% filter(RNA_library=="poly-A") 
#total RNAseq samples in histology
nrow(clinical_rna_Polya)

# check sample names of FPKM file in newly sequenced data and new release histologies
newSample<-colnames(newPolyaFPKM[,-1]) 
fpkmSampleCheck1<-setdiff(newSample , clinical_rna_Polya$Kids_First_Biospecimen_ID)
fpkmSampleCheck2<-setdiff( clinical_rna_Polya$Kids_First_Biospecimen_ID,newSample)


# check sample names of TPM file in newly sequenced data and new release histologies
newSample<-colnames(newPolyaTPM[,-1]) 
tpmSampleCheck1<-setdiff( newSample , clinical_rna_Polya$Kids_First_Biospecimen_ID)
tpmSampleCheck2<-setdiff( clinical_rna_Polya$Kids_First_Biospecimen_ID,newSample)

# check sample names of count file in newly sequenced data and new release histologies
newSample<-colnames(newPolyaCount[,-1]) 
countSampleCheck1<-setdiff( newSample , clinical_rna_Polya$Kids_First_Biospecimen_ID)
countSampleCheck2<-setdiff( clinical_rna_Polya$Kids_First_Biospecimen_ID,newSample)

if (all(is_empty(c(countSampleCheck1,tpmSampleCheck1,fpkmSampleCheck1,countSampleCheck2,tpmSampleCheck2,fpkmSampleCheck2)))){
  print ("New samples in rsem files completely overlap with RNAseq files in new histology file")
  SampleCountCheck<-TRUE
} else{
  stop ("Sample count inaccurate! Check Sample in new and old datasets")
  SampleCountCheck<-FALSE
}
```


```{r Check colnames for polya biospecimen should have "gene_id" for gene files }

# check col names of FPKM file of v18 in newly merged file and expected output
oldSample<-colnames(oldPolyaFPKM) 
newSample<-colnames(newPolyaFPKM) 
fpkmColCheck1<-setdiff(newSample , c("gene_id",clinical_rna_Polya$Kids_First_Biospecimen_ID))
fpkmColCheck2<-setdiff( c("gene_id",clinical_rna_Polya$Kids_First_Biospecimen_ID),newSample)

# check col names of count file of v18 in newly merged file and expected output
oldSample<-colnames(oldPolyaCount) 
newSample<-colnames(newPolyaCount) 
countColCheck1<-setdiff(newSample , c("gene_id",clinical_rna_Polya$Kids_First_Biospecimen_ID))
countColCheck2<-setdiff( c("gene_id",clinical_rna_Polya$Kids_First_Biospecimen_ID),newSample)

# check col names of TPM file of v18 in newly merged file and expected output
oldSample<-colnames(oldPolyaTPM) 
newSample<-colnames(newPolyaTPM) 
tpmColCheck1<-setdiff(newSample , c("gene_id",clinical_rna_Polya$Kids_First_Biospecimen_ID))
tpmColCheck2<-setdiff( c("gene_id",clinical_rna_Polya$Kids_First_Biospecimen_ID),newSample)

if (all(is_empty(c(countColCheck1,tpmColCheck1,fpkmColCheck1,countColCheck2,tpmColCheck2,fpkmColCheck2)))){
  print ("Combined samples in datasets completely overlap with RNAseq files in histology file")
  SampleCountCheck<-TRUE
} else{
  stop ("Colnames incorrect! Check colnames in new and old datasets")
  SampleCountCheck<-FALSE
}

```


```{r Check order of genes in old polya and new stranded dataset }
# only if number of gene_ids is true use == and all() to compare each element to the same position in the other vector

# check gene order kallisto polya and new kallisto dataset
if(length(newStrandedFPKM$gene_id)== length(oldPolyaFPKM$gene_id)){
  geneOrderCheckFPKM<-all(newStrandedFPKM$gene_id==oldPolyaFPKM$gene_id)
} else{
  geneOrderCheckKallisto<-FALSE
}

# check gene order polya and new TPM dataset
if(length(newStrandedTPM$gene_id)== length(oldPolyaTPM$gene_id)){
  geneOrderCheckTPM<-all(newStrandedTPM$gene_id==oldPolyaTPM$gene_id)
}else{
  geneOrderCheckTPM<-FALSE
}

# check gene order polya and new count dataset
if(length(newStrandedCount$gene_id)== length(oldPolyaCount$gene_id)){
  geneOrderCheckCount<-all(newStrandedCount$gene_id==oldPolyaCount$gene_id)
} else {
  geneOrderCheckCount<-FALSE
}

if (all(c(geneOrderCheckCount,geneOrderCheckFPKM,geneOrderCheckTPM))){
  print ("Same order of genes in both datasets ")
  geneOrderCheck<-TRUE
} else{
  stop ("Different genes order found! Check genes in new and old datasets")
  geneOrderCheck<-FALSE
}
```



```{r  Check all Checks and save RDS}
if(all(geneCountCheck,geneOrderCheck,SampleCountCheck)){
  print("Checks PASSED; coping files to release folder")
  saveRDS(newPolyaCount,file.path(outputFolder,params$polyaCount))
  saveRDS(newPolyaTPM,file.path(outputFolder,params$polyaTPM))
  saveRDS(newPolyaFPKM,file.path(outputFolder,params$polyaFPKM))
  
  saveRDS(newStrandedCount,file.path(outputFolder,params$strandedCount))
  saveRDS(newStrandedTPM,file.path(outputFolder,params$strandedTPM))
  saveRDS(newStrandedFPKM,file.path(outputFolder,params$strandedFPKM))
  
} else{
  stop("Checks FAILED. Please REVIEW")
}
```